---
title: "Replicate Analysis"
author: "Zac Driscoll"
date: "9/2/2021"
output: 
  html_document:
    code_folding: hide
---
## {.tabset}

```{r setup, include=FALSE}
library(mmsd.wq)
library(tidyverse)
library(sf)
library(leaflet)

knitr::opts_chunk$set(
message = FALSE, warning = FALSE
)

load("all_data.RData")

```


### General Analysis
<h3> Background </h3>
<ul type="square">
    <li>6 years of data was downloaded from LIMS (2015 - Present) </li>
    <li>Percent Exceedance was calculated by year, parameter, waterbody and site. </li>
    <li>Data were graphed and mapped in subsequent tabs. 
</ul> 

<h3> Major Conclusions </h3>
<ul type="square">
    <li>Samples in 2021 have a similar percent exceedance compared to previous years. </li>
    <li> Chl-a, TSS, and VSS have substantially more exceedances compared to other parameters </li>
    <li> The number of chl-a exceedances have dropped over time </li>
    <li> Cedar Creek has a high percentage of VSS and TSS exceedances compared to 
    other sites.  Root River has the most Chl-a exceedances. </li>
    <li> The Kinnickinnic River has more exceedances compared to other waterbodies. It
    also has more exceedances in 2021 compared to previous years (~13% of samples).</li>
    <li> LC-02, RI-33,and HC-1 have the most exceedances compared to all sites (>15% samples collected)
</ul> 


<h3> Percent Exceedances over time </h3>
The median percent exceedance by year is 6.4.  This is equal to the percent 
currently observed in 2021.
```{r,  fig.width = 10, fig.height = 7}

p <-data_yr[[2]] %>%
  filter(perc_exceeds != 0) %>%
  ggplot() +
  geom_bar(aes(x = year,y = perc_exceeds),
           stat = "identity") +
  ylab("Percent Exceedane")

plotly::ggplotly(p)

```
<br>

<h3> Percentage of Exceedances by Parameter </h3>
The three parameters that have the highest percent exceedances are chlorophyll a
VSS, and TSS. 
```{r,  fig.width = 10, fig.height = 7}

p <-data_param[[2]] %>%
  filter(perc_exceeds != 0) %>%
  left_join(param_lookup_table %>%
              select(parameter,label_name)) %>%
  select(-parameter) %>%
  rename(parameter = label_name) %>% 
  mutate(parameter = fct_reorder(parameter,perc_exceeds,max)) %>%
  ggplot() +
  geom_bar(aes(x = parameter,y = perc_exceeds),
           stat = "identity") +
  ylab("Perent Exceedance") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotly::ggplotly(p)

```
<br>
<h3> Percentage of Exceedances by Waterbody </h3>
The Kinnickinnic River has the highest percentage of samples exceeding
the replicate standard compared to all watersheds. 
```{r,  fig.width = 10, fig.height = 7}

p <-data_wtb[[2]] %>%
  filter(perc_exceeds != 0) %>%
  mutate(water_body = fct_reorder(water_body,perc_exceeds,max)) %>%
  ggplot() +
  geom_bar(aes(x = water_body,y = perc_exceeds),
           stat = "identity") +
  ylab("Perent Exceedance") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotly::ggplotly(p)

```

<br>
<h3> Percentage of Exceedances by Site </h3>
Use the slider bar to filter sites by percent exceedance. 
```{r,  fig.width = 10, fig.height = 7}

site_perc <- 
  data_site[[2]] %>%
   left_join(mmsd_wq_sites %>%
              select(sitecode,
                     latitude,
                     longitude)) %>%
  st_as_sf(coords = c("longitude","latitude"),crs = 4326) 

site_perc_shared <- crosstalk::SharedData$new(site_perc)

crosstalk::filter_slider("range_exceed2", "Percent Exceedance Range", 
              site_perc_shared, ~perc_exceeds, width = "100%")

pal <- colorNumeric(
  palette = colorRampPalette(c('green','yellow','orange','red'))(100), 
  domain = site_perc$perc_exceeds)

leaflet()%>%
  leaflet::addPolylines(
    data = mke_rivers,
    color = "blue",
    popup = ~ paste(" River: ", ROW_NAME)) %>% 
  addCircles(data = site_perc_shared,
              color = ~pal(perc_exceeds),
              fillColor = ~pal(perc_exceeds),
             fillOpacity = 0.85,
             radius = 200,
              popup = ~ paste(" Site: ",sitecode,"<br>",
                              " Exceedances: ",exceeds,"<br>",
                              " Samples: ",samples,"<br>",
                              " Percent Exceedance: ",perc_exceeds,"%",
                              sep =""))%>%
  addProviderTiles(providers$Esri.WorldImagery)%>%
  addLegend("bottomright", pal  = pal, values = site_perc$perc_exceeds,
            title = "Percent Exceedance",
            labFormat = labelFormat(suffix = "%"),
            opacity = 1
  )


```

### Parameter:Year

<br>
<h3> Figure 2: Heatmap - Parameter:Year </h3>
A heatmap showing the percentage of samples that exceeded the replicate standard
by year. Red cells indicate a higher percentage of samples exceeding the standard.
The number of chlorophyll exceedances has steadily dropped over the last few years.
<br>
```{r, fig.width = 10, fig.height = 7, warning= FALSE}

#heat map - param wtb
z <- data_yr_param[[2]] %>%
  filter(!parameter %in% c("hardness", 
                           "total_inorganic_carbon_2",
                           "nitrate_nitrogen_auto_chemistry")) %>%
  left_join(param_lookup_table %>%
              select(parameter,label_name)) %>%
  group_by(label_name) %>%
  mutate(sum_exceeds = sum(perc_exceeds))  %>%
  ungroup() %>%
  mutate(label_name = forcats::fct_reorder(label_name, sum_exceeds, max)) %>%
  rename(Percent_Exceedance = perc_exceeds) 

p <-  ggplot(data = z) +
  geom_tile(aes(
    x = year,
    y = label_name,
    text = samples,
    labels = exceeds,
    fill = Percent_Exceedance
  )) +
  geom_text(aes(
    x = year,
    y = label_name,
    label = paste0(Percent_Exceedance, "%")
  ),
  size = 3) +
  scale_fill_gradient(low = "#d6e4ff", high = "red") +
  ylab("")+
  xlab("")

plotly::ggplotly(p)

```
<br>
<h3> Figure 3: Time series</h3>
Select a parameter to see the percentage of sample exceeding the replicate
exceedance over time. <br>
```{r, echo = FALSE, warning = FALSE}

data_shared <- data_yr_param[[2]] %>%
  left_join(param_lookup_table %>%
              select(parameter,label_name))

exceed_shared <- crosstalk::SharedData$new(data_shared)

  crosstalk::filter_select(
    "oc_seasonal_param",
    "Select a Parameter:",
    multiple = TRUE,
    sharedData = exceed_shared,
    ~ label_name
  )

plotly::plot_ly(
  data = exceed_shared,
  y = ~ perc_exceeds,
  x = ~ year,
  type = 'scatter',
  mode = 'line',
  color = ~label_name
)

```

### Waterbody:Year

<h3> Figure 5: Waterbdy:Year Heatmap </h3>
A heatmap showing the percentage of samples that exceeded the replicate standard
by waterbody Red cells indicate a higher percentage of samples exceeding the standard. 
The Kinnickinnic River has more exceedances this year compared to previous years.

<br>
```{r, fig.width = 10, fig.height = 7, warning= FALSE}
p <- data_yr_wtb[[2]] %>%
  group_by(water_body) %>%
  mutate(sum_exceeds = sum(perc_exceeds))  %>%
  ungroup() %>%
  mutate(water_body = forcats::fct_reorder(water_body, sum_exceeds, max)) %>%
  rename(Percent_Exceedance = perc_exceeds) %>%
  ggplot() +
  geom_tile(aes(
    x = year,
    y = water_body,
    text = samples,
    labels = exceeds,
    fill = Percent_Exceedance
  )) +
  geom_text(aes(
    x = year,
    y = water_body,
    label = paste0(Percent_Exceedance, "%")
  ),
  size = 3) +
  scale_fill_gradient(low = "#d6e4ff", high = "red") +
  ylab("")+
  xlab("")

plotly::ggplotly(p)

```

<br>
<h3> Figure 6: Water body exceedance time series</h3>
Select a water body to see the percentage of sample exceeding the replicate
exceedance over time. <br>
```{r, echo = FALSE, warning = FALSE}

data_shared <- data_yr_wtb[[2]] 

exceed_shared <- crosstalk::SharedData$new(data_shared)

  crosstalk::filter_select(
    "yr_wtb",
    "Select a WaterBody:",
    multiple = TRUE,
    sharedData = exceed_shared,
    ~ water_body
  )

plotly::plot_ly(
  data = exceed_shared,
  y = ~ perc_exceeds,
  x = ~ year,
  type = 'scatter',
  mode = 'line',
  color = ~water_body
)

```

### Waterbody:Parameter

<h3> Figure 8: Waterbody:Parameter Heatmap </h3>
A heatmap showing the percentage of samples that exceeded the replicate standard
by waterbody and parameter. Red cells indicate a higher percentage of 
samples exceeding the standard. Cedar Creek has the highest observed exceedances
for both TSS and VSS. <br>
```{r, fig.width = 12, fig.height = 10, warning= FALSE}

#heat map - param wtb
p <- data_param_wtb[[2]] %>%
  filter(!parameter %in% c("hardness", 
                           "total_inorganic_carbon_2",
                           "nitrate_nitrogen_auto_chemistry")) %>%
  left_join(param_lookup_table %>%
              select(parameter,label_name)) %>%
  group_by(label_name) %>%
  mutate(sum_exceeds = sum(perc_exceeds))  %>%
  ungroup() %>%
  mutate(label_name = forcats::fct_reorder(label_name, sum_exceeds, max)) %>%
  rename(Percent_Exceedance = perc_exceeds) %>%
  ggplot() +
  geom_tile(aes(
    y = label_name,
    x = water_body,
    text = samples,
    labels = exceeds,
    fill = Percent_Exceedance
  )) +
  geom_text(aes(
    y = label_name,
    x = water_body,
    label = paste0(Percent_Exceedance, "%")
  ),
  size = 3) +
  scale_fill_gradient(low = "#d6e4ff", high = "red") +
  ylab("") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotly::ggplotly(p)

```

<br>
<h3> Figure 9: Water Body:Parameter Scater Plot</h3>
Select a parameter to see the percentage of sample exceeding the replicate
exceedance by water body  <br>
```{r, echo = FALSE, warning = FALSE,fig.width= 10, fig.height=10}

data_shared <- data_param_wtb[[2]] 

exceed_shared <- crosstalk::SharedData$new(data_shared)

  crosstalk::filter_select(
    "param_wtb",
    "Select a Parameter:",
    multiple = TRUE,
    sharedData = exceed_shared,
    ~ parameter
  )

plotly::plot_ly(
  data = exceed_shared,
  y = ~ perc_exceeds,
  x = ~ water_body,
  type = 'scatter',
  mode = 'line',
  color = ~parameter
)

```

### Exceedance Site Map

<h3> Page Summary </h3>
The  percentage of samples exceeding the replicate standard were 
calculated by site, parameter, and depth. Select a parameter and depth
to see how the percent exceedances vary by site. <br>

```{r, echo = FALSE, fig.width=10,fig.height = 10}

site_exceeds <- data_param_site[[2]] %>%
  left_join(mmsd_wq_sites %>%
              select(sitecode,
                     latitude,
                     longitude)) %>%
  st_as_sf(coords = c("longitude","latitude"),crs = 4326) 

site_shared <-  crosstalk::SharedData$new(site_exceeds)
crosstalk::bscols(
  crosstalk::filter_select(
    "param_map",
    "Select a Parameter:",
    multiple = FALSE,
    sharedData = site_shared,
    ~parameter
  ),
   crosstalk::filter_select(
    "depth_map",
    "Select a Depth:",
    multiple = FALSE,
    sharedData = site_shared,
    ~depth
  )
)

crosstalk::filter_slider("range_exceed", "Percent Exceedance", 
              site_shared, ~perc_exceeds, width = "100%")

pal <- colorNumeric(
  palette = colorRampPalette(c('green','yellow','orange','red'))(100), 
  domain = site_shared$perc_exceeds)

leaflet()%>%
  leaflet::addPolylines(
    data = mke_rivers,
    color = "blue",
    popup = ~ paste(" River: ", ROW_NAME)) %>% 
  addCircles(data = site_shared,
              color = ~pal(perc_exceeds),
              fillColor = ~pal(perc_exceeds),
             fillOpacity = 0.85,
             radius = 200,
              popup = ~ paste(" Site: ",sitecode,"<br>",
                              " Exceedances: ",exceeds,"<br>",
                              " Samples: ",samples,"<br>",
                              " Percent Exceedance: ",perc_exceeds,"%",
                              sep =""))%>%
  addProviderTiles(providers$Esri.WorldImagery)%>%
  addLegend("bottomright", pal  = pal, values = site_exceeds$perc_exceeds,
            title = "Percent Exceedance",
            labFormat = labelFormat(suffix = "%"),
            opacity = 1
  )
```


